% First of all, include a document class and options
\documentclass[english,bachelor]{diploma}
% There are some additional packages
\usepackage[autostyle=true]{csquotes} % enhanced support for quotation marks, support for biblatex package
\usepackage[backend=biber, style=iso-numeric, alldates=iso]{biblatex} % bibliography
\usepackage{dcolumn} % numeric column type
% \usepackage{subfig} % subtables and subfigures
\usepackage[julia]{diplomalst}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{hyperref}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subcaption}
\usepackage{amssymb}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% my own commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand\norm[1]{\lVert#1\rVert}
\definecolor{darkred}{RGB}{139,0,0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% Next, enter data for leading pages
\ThesisAuthor{Jonáš Koziorek}

\ThesisSupervisor{prof. RNDr. Marek Lampart, Ph.D.}

\CzechThesisTitle{Algoritmická detekce intermitence typu I}

\EnglishThesisTitle{Algorithmic detection of type-I intermittency}

\SubmissionYear{2024}

\ThesisAssignmentFileName{ThesisAssignment.pdf}

\Acknowledgement{
    I would like to thank prof. RNDr. Marek Lampart, Ph.D. who guided me through the process of writing this thesis. 
    I would also like to thank to all my teachers at VŠB-TUO who taught me so much in such a short time.
}

\CzechAbstract{
    Cílem práce je vyvinout algoritmus k detekci intermitence typu I v diskrétních dynamických systémech.
    Intermitence v dynamickém systému je jev, kdy se stav systému mění mezi zdánlivě stabilním a chaotickým.
    Práce se zaměřuje na intermitenci typu I, která souvisí se sedlovými bifurkacemi.
    Nejprve je představena teorie diskrétních dynamických systémů a intermitence typu I.
    Dále je vysvětleno, proč má detekce intermitence typu I smysl.
    Kombinací několika algoritmů od jiných autorů je detekce dosažena pro logistické zobrazení.
    Výsledky detekce jsou prezentovány graficky pomocí vybarvení bifurkačního diagramu.
    Nakonec je zmíněna implementace v programovacím jazyce Julia.
    }

\CzechKeywords{
    intermitence typu I; diskrétní dynamické systémy; logistické zobrazení; pevný bod; sedlová bifurkace; průměrná délka laminární fáze; programovací jazyk Julia; bifurkační diagram
    }

\EnglishAbstract{
    The goal of the thesis is to develop an algorithm for detection of type-I intermittency in discrete dynamical systems.
    Intermittency in a dynamical system is a phenomenon where the state of the system alternates between being seemingly stable and chaotic.
    The focus of the thesis is on the type-I intermittency which is associated with saddle-node bifurcations.
    Firstly, the theory of discrete dynamical systems and of type-I intermittency is introduced.
    Furthermore, the importance of detection of type-I intermittency is explained.
    By combining several algorithms of other authors, the detection is achieved for the Logistic map.
    The results of the detection are presented graphically through colorization of a bifurcation diagram.
    Lastly the implementation in Julia programming language is mentioned.
    }

\EnglishKeywords{
    type-I intermittency; discrete dynamical systems; Logistic map; fixed point; saddle-node bifurcation; average laminar phase length; Julia programming language; bifurcation diagram
    }

\AddAcronym{$\mathcal{L}_{r}$}{Logistic map}
\AddAcronym{$\mathcal{P}_{\varepsilon}$}{Pomeau-Manneville map}
\AddAcronym{DDS}{discrete dynamical system}
\AddAcronym{SPO}{stable periodic orbit}
\AddAcronym{UPO}{unstable periodic orbit}
\AddAcronym{DL}{Davidchack-Lai algorithm}
\AddAcronym{SD}{Schmelcher-Diakonos algorithm}
\AddAcronym{BWJ}{Bu-Wang-Jiang algorithm}
\AddAcronym{NLS}{Naive Local Search}
\AddAcronym{NGS}{Naive Global Search}
\AddAcronym{PSO}{Particle Swarm Optimization}
\AddAcronym{NLPSO}{Nested-Layer Particle Swarm Optimization}
\AddAcronym{LSGO}{Local Search using Global Optimization}

% Bibliography resources for BibLaTeX
\addbibresource{bibliography.bib}
\addbibresource{bibliography2.bib}

% New table column type for numeric values
\newcolumntype{d}[1]{D{.}{.}{#1}}

% Beginning of the document
\begin{document}

% Leading pages printing
\MakeTitlePages

% Do theses contain any figures? If so, print the list of figures and then move to the next page.
% If not, delete the next two macros.
\listoffigures
\clearpage

% Do theses contain any tables? If so, print the list of table and then move to the next page.
% If not, delete the next two macros.
% \listoftables
% \clearpage

% Your text starts here...
\input{Chapters/Introduction.tex}
\input{Chapters/Discrete Dynamical Systems.tex}
\input{Chapters/Type-I Intermittency.tex}
\input{Chapters/Intermittency Detection.tex}
\input{Chapters/Software.tex}
\input{Chapters/Conclusion.tex}

\printbibliography[heading=bibintoc]

\appendix
% \input{Chapters/Appending Complex Figures.tex}
% \input{Chapters/Appendix Codes.tex}

\end{document}